Meta-analysis

### 2019-11-01 03_Preliminary_analysis.Rmd
Moving preliminary analyses and plotting to chapter 3

### Yield saved from fungicide applications

The data plotted below are from the past field trials (2011 - 2017) described in the aforementioned table.
All data are a list of mean values from each treatment applied in each experiment.
Various factors have been studied in the previous experiments which may influence the following results. The following variables were disregarded in the plots below.

@@

```{r echo=FALSE, message=FALSE, warning=FALSE, include= FALSE}
# Import data
PM_MB_means <- import_data()
colnames(PM_MB_means)
class(PM_MB_means$year)
```

@@

>>>>

## Explore and visualise Data

```{r echo=FALSE, message=FALSE, warning=FALSE, include= FALSE}
# Import data
PM_MB_means <- import_data()

Various factors have been studied in the previous experiments which may influence the following results. However unless stated the following plots and analyses were made disregarding effects from other factors not mentioned.

@@

1. Type of Fungicide sprayed
2. In-crop rainfall
3. mungbean cultivar
4. Fungicide dose (we don't expect this to influence as it was only changes slightly)
5. Mean daily temperature in the weeks following spray application
6. Row spacing
7. Two pathogen species cause PM (new discovery)
8. irrigated vs non-irrigated trials
9. Number of fungicide sprays (edited)

### Fungicides

The data plotted below are from the past field trials (2011 - 2017) described in the aforementioned table.
All data are a list of mean values from each treatment applied in each experiment.

When investigating previous studies into powdery mildew fungicide efficacy on mungbean `r length(unique(PM_MB_means$trial_ref))` trials were found.
However, only eleven tebuconazole were used for the following meta-analysis, followed by the next most common fungicide, propiconazole. Sulphur was disregarded due to its inconsistency in efficacy.

```{r Fungicides}
PM_MB_means %>%
   group_by(fungicide_ai, trial_ref) %>%
   summarise() %>%
   count(sort = TRUE) %>%
   rename(Trials = n) %>%
   ggplot(aes(x = reorder(fungicide_ai, Trials), y = Trials)) +
   xlab("Fungicide active ingredient") +
   ylab("N Trials") +
   geom_col() +
   scale_fill_usq() +
   ggtitle(label = "Number of trials in which the\nspecified fungicide was used") +
   scale_colour_usq() +
   theme_usq() +
   coord_flip()
```

Tebuconazole and propiconazole might be able to be pooled as a triazole fungicide treatment. 
Amistar Xtra is a combination of azoxystrobin and cyproconazole.
Custodia is a combination of tebuconazole and azoxystrobin.
Amistar Xtra and Custodia might be able to be pooled given they both contain strobilurin and triazole, however, they contain differing dose ratios (inverted).

Perhaps best way forward is to do an analysis of only the triazoles. Then do another including azoxystrobin as a comparison.

Preliminary analysis of fungicides show there is a difference the effect of fungicide in proportion of yield saved.

#### Fungicide Doses

All trials that used tebuconazole used approximately the same dose.
Dose of the active ingredient ranged from 62.35&nbsp;g per hectare to 60&nbsp;g per hectare.

```{r tebu_dose}
PM_MB_means %>%
   filter(fungicide_ai == "tebuconazole") %>%
   select(trial_ref,
          year,
          location,
          first_sign_disease,
          dose_ai.ha,
          total_fungicide) %>%
   ggplot(aes(x = as.factor(dose_ai.ha))) +
   xlab("Dose (g ai/ha)") +
   ggtitle(label = "Total number of treatments for each respective tebuconazole dose") +
   geom_bar() +
   scale_fill_usq() +
   scale_colour_usq() +
   theme_usq()
```

### Row spacing

Some experiments were designed to investigate the effect of row spacing and plant density on powdery mildew disease.
The results showed that the row spacing had no statistically significant effect on powdery mildew, but narrower rows in most cases increased yield significantly. This finding has also been shown by Kerry McKenzie's work as well.
Eight trials used a row spacing of 0.75&nbsp;meters and tebuconazole as an active ingredient (AI).

```{r row_spacing.m}
PM_MB_means %>%
   filter(fungicide_ai == "tebuconazole" | fungicide_ai == "propiconazole") %>%
   group_by(fungicide_ai, row_spacing, trial_ref) %>%
   summarise() %>%
   count() %>%
   rename(Trials = n) %>%
   ggplot(aes(x = as.factor(row_spacing), y = Trials)) +
   xlab("Row Spacing (m)") +
   ylab("N Trials") +
   ggtitle(label = "Trial row spacing using tebuconazole") +
   geom_col(aes(fill = fungicide_ai),
            position = "dodge") +
   scale_fill_usq(name = "Fungicide AI") +
   theme_usq()

PM_MB_means %>%
   filter(fungicide_ai == "tebuconazole",
         row_spacing == 0.75) %>% 
   glimpse()
```

Were there any statistical difference between all row spacing or only some?
Can we pool certain row spacing that have no significant difference?
From the preliminary analysis, the graphs seem to imply that if there is a lower overall yield there is no effect of row spacing on yield.
However, if the average yield is more than approximately 0.6 - 1 t/ha then smaller row spacing has the potential to provide greater yield.

@@

### Host genotypes

Host genotype may need to be analysed to determine the effect.

```{r Host_genotype}
PM_MB_means %>%
   filter(fungicide_ai == "tebuconazole",
         row_spacing == 0.75) %>%
   group_by(host_genotype, trial_ref) %>%
   summarise() %>%
   count() %>%
   rename(Treatments = n) %>%
   ggplot(aes(x = host_genotype, y = Treatments)) +
   xlab("Cultivar") +
   ylab("N Trials") +
   ggtitle(label = "Cultivars used in tebuconazole trials with 0.75 m row spacing") +
   geom_col() +
   scale_fill_usq() +
   scale_colour_usq() +
   theme_usq()
```

In general the mungbean varieties have the following resistance to powdery mildew.

   - Berken: Highly susceptible

   - Crystal: Susceptible

   - Jade: Moderately susceptible

#### Genotype yield variability

Host genotype seems to contribute to the variation in yield. Berken shows higher yields compared to crystal and jade.
```{r yield_vol}
source("R/yield_volatility.r") #function to investigate the volatility in yields

yield_volatility(genotype_by_trial = FALSE, control_only = FALSE)
yield_volatility(genotype_by_trial = FALSE, control_only = TRUE)
yield_volatility(genotype_by_trial = FALSE, control_only = FALSE, location = "Hermitage")

yield_volatility("Crystal",control_only = TRUE)
yield_volatility("Crystal",control_only = FALSE)

yield_volatility("Jade",control_only = TRUE)
yield_volatility("Jade",control_only = FALSE, location = "Hermitage")

### Yield vs in-season rain

In crop rain could be a significant factor which contributes to the end of season grain yield. To investigate any relationship I will loop a linear model over time periods from 20 days prior to planting, to 100 days after planting.

#### Best start day for in-crop rainfall

Loop over the dates at the start of the season (to 90 days after planting) to find which date fits the best linear model for grain weight and in-crop rainfall. The code for this process was moved to be contained within it's own script and can be run as a job as it can take a long time. The output of the script is saved with the prefix `lmInSeasonRainfall` followed by the time windows the model is run. For example `lmInSeasonRainfall_20.40_50.80.csv` holds the model results for start days between 20 and 40 days after planting to end window dates between 50 and 80 days after planting.

```{r rainfall_sum_Sstart}
# The four data files were created from four scripts run as four jobs in Rstudio
# when the four jobs are finished they can be imported and combined into one data.frame
# source(Rainfall_x_cropYield_s-10.09_e30.65.R)
# source(Rainfall_x_cropYield_s-10.09_e65.100.R)
# source(Rainfall_x_cropYield_s10.30_e30.65.R)
# source(Rainfall_x_cropYield_s10.30_e65.10.R)

lm_rain1 <- read.csv("data/lmInSeasonRainfall_-10.09_30.64.csv",stringsAsFactors = FALSE)
lm_rain2 <- read.csv("data/lmInSeasonRainfall_-10.09_65.100.csv",stringsAsFactors = FALSE)
lm_rain3 <- read.csv("data/lmInSeasonRainfall_10.30_30.64.csv",stringsAsFactors = FALSE)
lm_rain4 <- read.csv("data/lmInSeasonRainfall_10.30_65.100.csv",stringsAsFactors = FALSE)

lm_rain <- rbind(lm_rain1,lm_rain2,lm_rain3,lm_rain4)

lm_rain[which(lm_rain$lm_pval == min(lm_rain$lm_pval)),]
lm_rain[which(lm_rain$lm_rsquared == max(lm_rain$lm_rsquared)),]
lm_rain[which(lm_rain$lm_adj_rsquared == max(lm_rain$lm_adj_rsquared)),]

ggplot(lm_rain, aes(x = start_day, y = end_day, z = log(lm_pval)))+
   geom_raster(aes(fill = log(lm_pval)), interpolate = TRUE)+
   geom_contour(bins = 15, colour = "white")+
   geom_abline(intercept = 0, slope = 1, colour = "red")+
   geom_point(data = lm_rain[which(lm_rain$lm_pval == min(lm_rain$lm_pval)),], aes(x = start_day, y = end_day), colour =" red")+
   scale_fill_distiller(palette ="RdBu", direction = -1)

ggplot(lm_rain, aes(x = start_day, y = end_day, z = lm_rsquared))+
   geom_raster(aes(fill = lm_rsquared), interpolate = TRUE)+
   geom_contour(bins = 15, colour = "white")+
   geom_abline(intercept = 0, slope = 1, colour = "red")+
   geom_point(data = lm_rain[which(lm_rain$lm_rsquared == max(lm_rain$lm_rsquared)),], aes(x = start_day, y = end_day), colour =" red")+
   scale_fill_distiller(palette ="RdBu")

ggplot(lm_rain, aes(x = start_day, y = end_day, z = lm_adj_rsquared))+
   geom_raster(aes(fill = lm_adj_rsquared), interpolate = TRUE)+
   geom_contour(bins = 15, colour = "white")+
   geom_abline(intercept = 0, slope = 1, colour = "red")+
   geom_point(data = lm_rain[which(lm_rain$lm_rsquared == max(lm_rain$lm_rsquared)),], aes(x = start_day, y = end_day), colour =" red")+
   scale_fill_distiller(palette ="RdBu")

The approximate period when in-crop rainfall impacts the most on the mean harvest is between 11 - 88 days after planting.

### Season range and first incidence

```{r first_incidence}
# Here I want to build a plot of horizontal lines indicating the season length and
# place a dot on the line to indicating the day of the year first sign was observed for that trial

season_dates <- PM_MB_means %>%
   group_by(trial_ref)%>%
   summarise(Planting_date = unique(planting_date),
             Harvest_date = unique(harvest_date),
             First_sign_PM = unique(first_sign_disease),
             season_length = as.integer(as.Date(unique(harvest_date)) - as.Date(unique(planting_date)))
             )

# function to extract date from the year
# format should be Universal time format
source("R/year2day.r")

season_dates$Planting_day <- unlist(lapply(season_dates$Planting_date,year2day))
season_dates$Disease_day <- unlist(lapply(season_dates$First_sign_PM,year2day))
season_dates$Harvest_day <- season_dates$Planting_day + season_dates$season_length

season_dates[1,"Planting_day"] <- year2day(as.Date("2012-02-28")-7) # planted in december and emerged in February, Assuming due to lack of rain or recording error. Using 7 days prior to emergence day as planting day

first_day_month <- c(0, 31, 59, 90, 120, 151, 181, 211, 243, 273)
axis_labels_date <- c(format(Sys.Date() - year2day(Sys.Date()) + first_day_month[1], "%b-%d"),  # Jan
                         format(Sys.Date() - year2day(Sys.Date()) + first_day_month[2], "%b-%d"), # Feb
                         format(Sys.Date() - year2day(Sys.Date()) + first_day_month[3], "%b-%d"), # March
                         format(Sys.Date() - year2day(Sys.Date()) + first_day_month[4], "%b-%d"), # April
                         format(Sys.Date() - year2day(Sys.Date()) + first_day_month[5], "%b-%d"), # May
                         format(Sys.Date() - year2day(Sys.Date()) + first_day_month[6], "%b-%d"), # June
                         format(Sys.Date() - year2day(Sys.Date()) + first_day_month[7], "%b-%d"), # July
                         format(Sys.Date() - year2day(Sys.Date()) + first_day_month[8], "%b-%d"), # August
                         format(Sys.Date() - year2day(Sys.Date()) + first_day_month[9], "%b-%d"), # Sept
                         format(Sys.Date() - year2day(Sys.Date()) + first_day_month[10], "%b-%d") # October
                         )

# reorder the factors so they will appear in chronilogical order
# I used First sign because all trials have recorded this
season_dates$trial_ref <- factor(season_dates$trial_ref, levels = levels(season_dates$trial_ref)[rev(order(season_dates$First_sign_PM))])

season_dates %>%
   ggplot()+
   geom_pointrange(aes(x = trial_ref, y = Disease_day, ymin = Planting_day, ymax = Harvest_day))+
   scale_y_continuous(
      limits = c(0, max(season_dates$Planting_day + 
                           season_dates$season_length, 
                        na.rm = TRUE)),
      labels = axis_labels_date,
      breaks = first_day_month
   )+
   geom_point(aes(x = trial_ref, y = Disease_day))+
   coord_flip()+
   ggtitle("Season length and when powdery mildew was first spotted\nin each respective trial")
